# Iteration

## Introduction

 
```{r message=FALSE}
library("tidyverse")
library("stringr")
```
The package **microbenchmark** is used for timing code
```{r}
library("microbenchmark")
```


## For Loops

### Exercises

Write for loops to:

1. Compute the mean of every column in `mtcars`.
2. Determine the type of each column in `nycflights13::flights`.
3. Compute the number of unique values in each column of `iris`.
4. Generate 10 random normals for each of $\mu = -10$, 0, 10, and 100.

Think about the output, sequence, and body before you start writing the loop.

To compute the mean of every column in `mtcars`.
```{r}
output <- vector("double", ncol(mtcars))
names(output) <- names(mtcars)
for (i in names(mtcars)) {
  output[i] <- mean(mtcars[[i]])
}
output
```

Determine the type of each column in `nycflights13::flights`.
Note that we need to use a `list`, not a character vector, since the class can have multiple values.
```{r}
data("flights", package = "nycflights13")
output <- vector("list", ncol(flights))
names(output) <- names(flights)
for (i in names(flights)) {
  output[[i]] <- class(flights[[i]])
}
output
```


```{r}
data(iris)
iris_uniq <- vector("double", ncol(iris))
names(iris_uniq) <- names(iris)
for (i in names(iris)) {
  iris_uniq[i] <- length(unique(iris[[i]]))
}
iris_uniq
```

```{r}
# number to draw
n <- 10
# values of the mean
mu <- c(-10, 0, 10, 100)
normals <- vector("list", length(mu))
for (i in seq_along(normals)) {
  normals[[i]] <- rnorm(n, mean = mu[i])
}
normals
```

However, we don't need a `for` loop for this since `rnorm` recycles means.
```{r}
matrix(rnorm(n * length(mu), mean = mu), ncol = n)
```

2. Eliminate the for loop in each of the following examples by taking advantage of an existing function that works with vectors:

```{r}
out <- ""
for (x in letters) {
  out <- stringr::str_c(out, x)
}
out
```

`str_c` already works with vectors, so simply use `str_c` with the `collapse` argument to return a single string.
```{r}
stringr::str_c(letters, collapse = "")
```

For this I'm going to rename the variable `sd` to something different because `sd` is the name of the function we want to use.
```{r}
x <- sample(100)
sd. <- 0
for (i in seq_along(x)) {
  sd. <- sd. + (x[i] - mean(x)) ^ 2
}
sd. <- sqrt(sd. / (length(x) - 1))
sd.
```

We could simply use the `sd` function.
```{r}
sd(x)
```
Or if there was a need to use the equation (e.g. for pedagogical reasons), then
the functions `mean` and `sum` already work with vectors:
```{r}
sqrt(sum((x - mean(x)) ^ 2) / (length(x) - 1))
```


```{r}
x <- runif(100)
out <- vector("numeric", length(x))
out[1] <- x[1]
for (i in 2:length(x)) {
  out[i] <- out[i - 1] + x[i]
}
out
```
The code above is calculating a cumulative sum. Use the function `cumsum`
```{r}
all.equal(cumsum(x),out)
```

**Ex. 21.2.1.3**  Combine your function writing and for loop skills:

    1. Write a for loop that `prints()` the lyrics to the children's song 
       "Alice the camel".

    1. Convert the nursery rhyme "ten in the bed" to a function. Generalise 
       it to any number of people in any sleeping structure.

    1. Convert the song "99 bottles of beer on the wall" to a function.
       Generalise to any number of any vessel containing any liquid on 
       any surface.
    
I don't know what the deal is with Hadley and nursery rhymes.
Here's the lyrics for [Alice the Camel](http://www.kididdles.com/lyrics/a012.html)

We'll look from five to no humps, and print out a different last line if there are no humps. This uses `cat` instead of `print`, so it looks nicer.
```{r}
humps <- c("five", "four", "three", "two", "one", "no")
for (i in humps) {
  cat(str_c("Alice the camel has ", rep(i, 3), " humps.",
             collapse = "\n"), "\n")
  if (i == "no") {
    cat("Now Alice is a horse.\n")
  } else {
    cat("So go, Alice, go.\n")
  }
  cat("\n")
}
```
我写的：
```{r}
num <- c('five', 'four', 'three', 'two', 'one', 'no')
for(i in seq_along(num)){
  line <- rep(paste('Alice the camel has ', num[[i]], ' humps.\n', sep = ''), 3)
  cat(line)
  if(i <= 5){
      cat('So go, Alice, go.\n')
  } else {
      cat('Now Alice is a horse.\n')
  }
}
```
输出为：
```
Alice the camel has five humps.  
 Alice the camel has five humps.  
 Alice the camel has five humps.  
So go, Alice, go.  
Alice the camel has four humps.
 Alice the camel has four humps.  
 Alice the camel has four humps.  
So go, Alice, go.  
Alice the camel has three humps.  
 Alice the camel has three humps.  
 Alice the camel has three humps.  
So go, Alice, go.  
Alice the camel has two humps.  
 Alice the camel has two humps.  
 Alice the camel has two humps.  
So go, Alice, go.  
Alice the camel has one humps.  
 Alice the camel has one humps.  
 Alice the camel has one humps.  
So go, Alice, go.  
Alice the camel has no humps.  
 Alice the camel has no humps.  
 Alice the camel has no humps.  
Now Alice is a horse.  
```


The lyrics for [Ten in the Bed](http://supersimplelearning.com/songs/original-series/one/ten-in-the-bed/):
```{r}
numbers <- c("ten", "nine", "eight", "seven", "six", "five",
             "four", "three", "two", "one")
for (i in numbers) {
  cat(str_c("There were ", i, " in the bed\n"))
  cat("and the little one said\n")
  if (i == "one") {
    cat("I'm lonely...")
  } else {
    cat("Roll over, roll over\n")
    cat("So they all rolled over and one fell out.\n")
  }
  cat("\n")
}
```

For the bottles of beer, I define a helper function to correctly print the number of bottles.
```{r}
bottles <- function(i) {
  if (i > 2) {
   bottles <- str_c(i - 1, " bottles")
  } else if (i == 2) {
   bottles <- "1 bottle"
  } else {
   bottles <- "no more bottles"
  }
  bottles
}

beer_bottles <- function(n) {
  # should test whether n >= 1.
  for (i in seq(n, 1)) {
     cat(str_c(bottles(i), " of beer on the wall, ", bottles(i), " of beer.\n"))
     cat(str_c("Take one down and pass it around, ", bottles(i - 1),
                " of beer on the wall.\n\n"))
  }
  cat("No more bottles of beer on the wall, no more bottles of beer.\n")
  cat(str_c("Go to the store and buy some more, ", bottles(n), " of beer on the wall.\n"))
}
beer_bottles(3)
```


**Ex 21.2.1.4**  It's common to see for loops that don't preallocate the output and instead increase the length of a vector at each step:
    
```{r, eval = FALSE}
output <- vector("integer", 0)
for (i in seq_along(x)) {
  output <- c(output, lengths(x[[i]]))
}
output
```

I'll use the package **microbenchmark** to time this.
The `microbenchmark` function will run an R expression a number of times and time it.

Define a function that appends to an integer vector.
```{r}
add_to_vector <- function(n) {
  output <- vector("integer", 0)
  for (i in seq_len(n)) {
    output <- c(output, i)
  }
  output  
}
microbenchmark(add_to_vector(10000), times = 3)
```

And one that pre-allocates it.
```{r}
add_to_vector_2 <- function(n) {
  output <- vector("integer", n)
  for (i in seq_len(n)) {
    output[[i]] <- i
  }
  output
}
microbenchmark(add_to_vector_2(10000), times = 3)
```

The pre-allocated vector is about **100** times faster!
YMMV, but the longer the vector and the bigger the objects, the more that pre-allocation will outperform appending.


## For loop variations


### 

**Ex** Imagine you have a directory full of CSV files that you want to read in.
You have their paths in a vector, 
`files <- dir("data/", pattern = "\\.csv$", full.names = TRUE)`, and now
want to read each one with `read_csv()`. Write the for loop that will 
load them into a single data frame. 

I pre-allocate a list, read each file as data frame into an element in that list.
This creates a list of data frames.
I then use `bind_rows` to create a single data frame from the list of data frames.
```{r eval = FALSE}
df <- vector("list", length(files))
for (fname in seq_along(files)) {
  df[[i]] <- read_csv(files[[i]])
}
df <- bind_rows(df)
```


**Ex** What happens if you use `for (nm in names(x))` and `x` has no names?
What if only some of the elements are named? What if the names are
not unique?

Let's try it out and see what happens.

When there are no names for the vector, it does not run the code in the loop (it runs zero iterations of the loop):
```{r}
x <- 1:3
print(names(x))
for (nm in names(x)) {
  print(nm)
  print(x[[nm]])
}
```
Note that the length of `NULL` is zero:
```{r}
length(NULL)
```

If there only some names, then we get an error if we try to access an element without a name. 
However, oddly, `nm == ""` when there is no name.
```{r}
x <- c(a = 1, 2, c = 3)
names(x)
```
```{r error = TRUE}
for (nm in names(x)) {
  print(nm)
  print(x[[nm]])
}
```

Finally, if there are duplicate names, then `x[[nm]]` will give the *first* element with that name.
There is no way to access elements with duplicate names.
```{r}
x <- c(a = 1, a = 2, c = 3)
names(x)
```
```{r}
for (nm in names(x)) {
  print(nm)
  print(x[[nm]])
}
```


**Ex**  Write a function that prints the mean of each numeric column in a data 
frame, along with its name. For example, `show_mean(iris)` would print:
    
```{r, eval = FALSE}
show_mean(iris)
#> Sepal.Length: 5.84
#> Sepal.Width:  3.06
#> Petal.Length: 3.76
#> Petal.Width:  1.20
```
    
(Extra challenge: what function did I use to make sure that the numbers
lined up nicely, even though the variable names had different lengths?)

There may be other functions to do this, but I'll use `str_pad`, and `str_length` to ensure that the space given to the variable names is the same.
I messed around with the options to `format` until I got two digits .
```{r}
show_mean <- function(df, digits = 2) {
  # Get max length of any variable in the dataset
  maxstr <- max(str_length(names(df)))
  for (nm in names(df)) {
    if (is.numeric(df[[nm]])) {
      cat(str_c(str_pad(str_c(nm, ":"), maxstr + 1L, side = "right"),
                format(mean(df[[nm]]), digits = digits, nsmall = digits),
                sep = " "),
          "\n")
    }
  
